---
title: "INLA for ECOFUNC"
output: word_document
---

```{r setup, include=FALSE}
require("knitr")
knitr::opts_chunk$set(echo = TRUE)
source("/home/zofie.cimburova/git/oekokart/ecofunc_helpful_functions.R")
opts_knit$set(root.dir = "/home/zofie.cimburova/ECOFUNC/DATA/OBSERVATIONS")
```

```{r, include=FALSE}
library(ggplot2)
```

# 1. Load the data
470 501 observations are exported from GRASS. For each observation point, its x- and y- coordinate, row and column number, total number of forest pixels, total number of forest and open land pixels and median values of environmental explanatory variables are stored.

```{r, message=FALSE}
input_data <- read.csv("/home/zofie.cimburova/ECOFUNC/DATA/OBSERVATIONS/observations_fenoscandia.csv")
```

```{r, include=FALSE}
# change names to shorter versions
names(input_data)[names(input_data)=="dem_10m_nosefi_float_aspect_sin"] <- "aspect_sin"
names(input_data)[names(input_data)=="dem_10m_nosefi_float_aspect_cos"] <- "aspect_cos"
names(input_data)[names(input_data)=="dem_10m_nosefi_float_slope"] <- "slope"
names(input_data)[names(input_data)=="dem_10m_nosefi_float_profc"] <- "curvature"
names(input_data)[names(input_data)=="dem_10m_topex_e"] <- "topex_e"
names(input_data)[names(input_data)=="dem_10m_topex_n"] <- "topex_n"
names(input_data)[names(input_data)=="dem_10m_topex_ne"] <- "topex_ne"
names(input_data)[names(input_data)=="dem_10m_topex_nw"] <- "topex_nw"
names(input_data)[names(input_data)=="dem_10m_topex_s"] <- "topex_s"
names(input_data)[names(input_data)=="dem_10m_topex_se"] <- "topex_se"
names(input_data)[names(input_data)=="dem_10m_topex_sw"] <- "topex_sw"
names(input_data)[names(input_data)=="dem_10m_topex_w"] <- "topex_w"
names(input_data)[names(input_data)=="dem_10m_topex_exposure"] <- "exposure"
names(input_data)[names(input_data)=="dem_tpi_250_50m"] <- "tpi_250"
names(input_data)[names(input_data)=="dem_tpi_500_50m"] <- "tpi_500"
names(input_data)[names(input_data)=="dem_tpi_1000_50m"] <- "tpi_1000"
names(input_data)[names(input_data)=="dem_tpi_2500_50m"] <- "tpi_2500"
names(input_data)[names(input_data)=="dem_tpi_5000_50m"] <- "tpi_5000"
names(input_data)[names(input_data)=="dem_tpi_10000_50m"] <- "tpi_10000"
names(input_data)[names(input_data)=="dem_10m_nosefi_tri"] <- "tri"
names(input_data)[names(input_data)=="solar_radiation_10m_april"] <- "solrad_apr"
names(input_data)[names(input_data)=="solar_radiation_10m_autumn"] <- "solrad_autumn"
names(input_data)[names(input_data)=="solar_radiation_10m_january"] <- "solrad_jan"
names(input_data)[names(input_data)=="solar_radiation_10m_july"] <- "solrad_jul"
names(input_data)[names(input_data)=="solar_radiation_10m_october"] <- "solrad_oct"
names(input_data)[names(input_data)=="solar_radiation_10m_spring"] <- "solrad_spring"
names(input_data)[names(input_data)=="solar_radiation_10m_summer"] <- "solrad_summer"
names(input_data)[names(input_data)=="solar_radiation_10m_winter"] <- "solrad_winter"
names(input_data)[names(input_data)=="solar_radiation_10m_year"] <- "solrad_year"
names(input_data)[names(input_data)=="bio01_eurolst_10m"] <- "bio01"
names(input_data)[names(input_data)=="bio02_eurolst_10m"] <- "bio02"
names(input_data)[names(input_data)=="bio10_eurolst_10m"] <- "bio10"
names(input_data)[names(input_data)=="bio11_eurolst_10m"] <- "bio11"
names(input_data)[names(input_data)=="bio12_eurolst_10m"] <- "bio12"
names(input_data)[names(input_data)=="bio15_eurolst_10m"] <- "bio15"
names(input_data)[names(input_data)=="bio18_eurolst_10m"] <- "bio18"
names(input_data)[names(input_data)=="bio02_eurolst_10m"] <- "bio02"
names(input_data)[names(input_data)=="bio19_eurolst_10m"] <- "bio19"
names(input_data)[names(input_data)=="sea_distance_10m"] <- "sea"
names(input_data)[names(input_data)=="sea_open_distance_50m"] <- "sea_open"
names(input_data)[names(input_data)=="water_distance_50m"] <- "water"
names(input_data)[names(input_data)=="latitude_10m"] <- "latitude"
names(input_data)[names(input_data)=="longitude_10m"] <- "longitude"

pred.names <- names(input_data)[c(6:8,17:42,45:48)]
```
```{r, message=FALSE}
input_data$lc_proportion <- input_data$nsuccess/input_data$ntrial
```

```{r, echo=FALSE}
head(input_data[,pred.names])
```


# 2. some filtering
All incomplete observations are removed, i.e. each observation point contains full data

```{r, message=FALSE}
input_data_na_rm <- input_data[complete.cases(input_data),]
```
```{r, echo=FALSE}
ggplot() + geom_tile(data=input_data_na_rm, aes(x=x, y=y, fill=exposure)) + coord_equal()
```


# 3. assign node ID for INLA
In order for INLA to work, each observation point is assigned an ID which determines its position in raster. Indexes are assigned vertically from top left corner to bottom right corner.
```{r, message=FALSE}
require(INLA)

nrows=1808
ncols=1414

input_data_na_rm$ID <- NA
input_data_na_rm$ID = inla.lattice2node(input_data_na_rm$row, 
                                        input_data_na_rm$col, 
                                        nrows, ncols)
```




# 4. standardize data
For logistic regression predictors need to be standardized to zero mean and standard deviation equal to one, i.e. column mean is subtracted and the result is divided by standard deviation. 
Column means and standard deviations are kept in order to be utilized in prediction.

```{r, message=FALSE}
library(robustHD)

input_data_na_rm.scaled <- input_data_na_rm
input_data_na_rm.scaled[,pred.names] <- standardize(input_data_na_rm[,pred.names])

input_data_na_rm.means <- colMeans(input_data_na_rm[,pred.names])
input_data_na_rm.stddev <- apply(input_data_na_rm[,pred.names], 2, sd)
```


# 5. exploratory data analysis
## Correlation coefficient of response~predictor
First, correlations between response and predictors are explored.

```{r, message=FALSE}
data.cor.coef <- data.frame(cor.coef=double())
i = 1
for (predictor in pred.names) {
  data.cor.coef[i,] <- cor(input_data_na_rm.scaled$lc_proportion, input_data_na_rm.scaled[,predictor])
  row.names(data.cor.coef)[i]<- predictor
  i = i+1
}
```
```{r, insert=FALSE}
data.cor.coef$predictor <- rownames(data.cor.coef)
data.cor.coef <- data.cor.coef[order(data.cor.coef$cor.coef),] 
data.cor.coef$predictor <- factor(data.cor.coef$predictor, levels = data.cor.coef$predictor[order(data.cor.coef$cor.coef)])
```

```{r, echo=FALSE}
ggplot(data.cor.coef, aes(predictor, cor.coef)) + geom_col() + coord_flip()
```


## Correlations between predictors
Correlations between predictors are checked and eliminated by three methods - first, correlation matrix is computed and pairs with correlation coefficient larger than 0.7 are explored. Second, full model is fitted with GLM, and predictors with variance inflation factor larger than 4 are exluded. Third, the data are projected on the first two principal components and predictors with similar directions are explored.

This procedure results in two sets of predictors - one keeping the indirectly influencing covariates - often correlated with similar directly influencing ones, but having better correlation coefficient with the response, the other one keeping only the directly influencing covariates. An example is the correlation between distance from sea, and bio02 temperature (approximation of continentality). Both of them represent a measure of continentality, but whilst distance from sea influences the probability of tree occurence indirectly (i.e. influences the climatic conditions), bio02 has a direct influence.

### Exluding indirectly influencing predictors
Following predictors were excluded:  

Prioritizing direct influences|Prioritizing indirect influences
----------------------------- |--------------------------------
height (*)                    |
aspect_sin                    |
aspect_cos                    |


### Correlation matrix
Correlation matrix was computed and pairs of covariates with **correlation coefficient larger than 0.7** were explored. Predictor with larger correlation coefficient with response was generally kept.


```{r, message=FALSE}
library(corrplot)

M.cor <- cor(input_data_na_rm.scaled[,pred.names])
M.cor.lim <- M.cor*(abs(M.cor)>=0.7)
```

```{r, echo=FALSE}
corrplot(M.cor.lim, method="circle",type="lower",order ="FPC")
```
Following predictors were excluded:  

Prioritizing direct influences|Prioritizing indirect influences
----------------------------- |--------------------------------
solrad_jul                    |solrad_jul 
solrad_spring                 |solrad_spring
solrad_summer                 |solrad_summer
solrad_year                   |solrad_year
solrad_apr                    |solrad_apr
solrad_jan (*oct)             |solrad_jan
solrad_winter                 |solrad_winter
solrad_autumn                 |solrad_autumn
tpi_2500                      |tpi_2500
tpi_250                       |tpi_250
tpi_500                       |tpi_500
tpi_5000                      |tpi_5000
bio12                         |bio12
bio18                         |bio18 
bio11                         |bio11  
tri                           |tri
bio01                         |bio01
sea_open                      |sea_open
sea                           |bio02


```{r, insert=FALSE}
pres.names.exclude.direct <- pred.names[!pred.names %in% 
                                 c("solrad_jul","solrad_spring","solrad_summer","solrad_year",
                                   "solrad_apr","solrad_winter","solrad_oct","solrad_autumn",
                                   "tpi_2500","tpi_250","tpi_500","tpi_5000",
                                   "bio12","bio18","bio11","tri","bio01","sea_open","sea",
                                   "longitude","latitude","aspect_sin","aspect_cos")]
pres.names.exclude.indirect <- pred.names[!pred.names %in% 
                                 c("solrad_jul","solrad_spring","solrad_summer","solrad_year",
                                   "solrad_apr","solrad_jan","solrad_winter","solrad_autumn",
                                   "tpi_2500","tpi_250","tpi_500","tpi_5000",
                                   "bio12","bio18","bio11","tri","bio01","sea_open","bio02")]

M.cor.direct <- cor(input_data_na_rm.scaled[,pres.names.exclude.direct])
M.cor.indirect <- cor(input_data_na_rm.scaled[,pres.names.exclude.indirect])

M.cor.direct.lim <- M.cor.direct*(abs(M.cor.direct)>=0.7)
M.cor.indirect.lim <- M.cor.indirect*(abs(M.cor.indirect)>=0.7)
```

```{r, echo=FALSE}
corrplot(M.cor.direct.lim, method="circle",type="lower",order ="FPC",title="Correlation matrix of prioritized direct influences")
corrplot(M.cor.indirect, method="circle",type="lower",order ="FPC",title="Correlation matrix of prioritized indirect influences")
```

###  VIF
A full logistic model (using all covariates left from previous step) is fitted with GLM, ommiting the spatial random effect. Predictors with **VIF larger than 4** were removed one-by-one using a backwards, stepwise procedure. 

Following predictors were excluded:  

Prioritizing direct influences|Prioritizing indirect influences
----------------------------- |--------------------------------
                              |solrad_oct


```{r, message=FALSE}
library(car)
formula.glm.direct <- lc_proportion ~ water + curvature + slope + tpi_1000 + bio10 + bio15 + 
                                      bio19 + tpi_10000 + exposure + bio02 + solrad_jan + height

formula.glm.indirect <- lc_proportion ~ water + curvature + slope + tpi_1000 + bio10 + bio15 + 
                                        bio19 + tpi_10000 + exposure + sea +
                                        height + aspect_sin + aspect_cos 

glm.direct <- glm(formula.glm.direct,
                  data=input_data_na_rm.scaled, 
                  family=binomial(link='logit'))

glm.indirect <- glm(formula.glm.indirect, 
                    data=input_data_na_rm.scaled, 
                    family=binomial(link='logit'))
vif(glm.direct)
vif(glm.indirect)
```

### PCA
Original variables were plotted in the 2D space of first two principal components. In case the cumulative propostion of variance explained by the first two principal components is high, the angles between vectors of original variables may be interpreted as correlations - small angles indicate the variables are positively correlated, an angle of 90 degrees indicates the variables are not correlated, and angles close to 180 degrees indicate the variables are negatively correlated.

For subset of variables prioritizing direct influences the cumulative proportion of variance explained by the first two components is 52 % and for subset of variables prioritizing indirect influences the cumulative proportion of variance explained by the first two components is 42 %. I assume this proportion is too low to enable interpreting the correlation between variables.

Following predictors were excluded:  

Prioritizing direct influences|Prioritizing indirect influences
----------------------------- |--------------------------------
tpi_1000                      |tpi_1000    


```{r, message=FALSE}

pca.direct <- princomp(input_data_na_rm.scaled[,c("water", "curvature", "slope", "bio10", "bio15" , "tpi_1000",
                                                  "bio19" , "tpi_10000" ,"exposure", "bio02","height","solrad_jan")], cor=TRUE) 

pca.indirect <- princomp(input_data_na_rm.scaled[,c("water", "curvature", "slope" , "bio10", "bio15" , "tpi_1000",
  "bio19" , "height" , "aspect_sin" ,"aspect_cos" , "tpi_10000" ,"exposure", "sea")], cor=TRUE) 

summary(pca.indirect) # variance accounted for
summary(pca.direct) # variance accounted for

#loadings(pca.direct) # pc loadings
```
```{r, echo=FALSE,fig.width=8,fig.height=8}
#biplot(pca.indirect,main = "Biplot from indirect subset") #this pulls out the variables on the first 2 axes
biplot(pca.direct,main = "Biplot from direct subset") #this pulls out the variables on the first 2 axes
```

## Final environmental variables used in modelling
Prioritizing direct influences|Prioritizing indirect influences
----------------------------- |--------------------------------
slope                         |slope
curvature                     |curvature
exposure                      |exposure
tpi_10000                     |tpi_10000
water                         |water
bio02                         |
bio10                         |bio10
bio15                         |bio15
bio19                         |bio19
                              |height
                              |aspect_sin
                              |aspect_cos
                              |sea

# 8. Transformation of covariates
Explanatory variables shall be normally distributed.

```{r, echo=FALSE}
library(rcompanion)
for (predictor in c("slope", "curvature", "exposure", "tpi_10000", "water", 
                    "bio02", "bio10", "bio15", "bio19", "height", "aspect_sin", 
                    "aspect_cos", "sea")) {
  hist <- plotNormalHistogram(input_data_na_rm[,predictor], main = predictor)
  print(hist)
}
```

Covariate   |Tranformation needed?|Name of transformed variable
----------- |---------------------|----------------------------
slope       |transform            |slope_t
curvature   |ok                   |curvature
exposure    |transform            |exposure_t
tpi_10000   |ok                   |tpi_10000
water       |transform            |water_t
bio02       |ok                   |bio02
bio10       |ok                   |bio10
bio15       |?                    |bio15
bio19       |transform            |bio19
height      |transform            |height_t
aspect_sin  |?                    |aspect_sin
aspect_cos  |?                    |aspect_cos
sea         |?                    |sea


## Slope - log transformation + 1
```{r, message=FALSE}
input_data_na_rm$slope_t <- log(input_data_na_rm$slope+1)
input_data_na_rm.scaled$slope_t <- standardize(input_data_na_rm$slope_t)

hist <- plotNormalHistogram(input_data_na_rm.scaled$slope_t, main = "Transformed slope")
print(hist)
```

## Exposure - log transformation + 1
Average value instead of median needs to be computed. Median exposure contatins too many zeros.
```{r, message=FALSE}
input_data_na_rm$exposure_t <- log(input_data_na_rm$exposure+1)
input_data_na_rm.scaled$exposure_t <- standardize(input_data_na_rm$exposure_t)

hist <- plotNormalHistogram(input_data_na_rm.scaled$exposure_t, main = "Transformed exposure")
print(hist)
```

## Water - cube root transformation
```{r, message=FALSE}
input_data_na_rm$water_t <- sign(input_data_na_rm$water) * abs(input_data_na_rm$water)^(1/3)
input_data_na_rm.scaled$water_t <- standardize(input_data_na_rm$water_t)

hist <- plotNormalHistogram(input_data_na_rm.scaled$water_t, main = "Transformed water proximity")
print(hist)
```

## Bio 19 - log transformation???
```{r, message=FALSE}
input_data_na_rm$bio19_t <- log(input_data_na_rm$bio19)
input_data_na_rm.scaled$bio19_t <- standardize(input_data_na_rm$bio19_t)

hist <- plotNormalHistogram(input_data_na_rm.scaled$bio19_t, main = "Transformed bio19")
print(hist)
```

## Height - Square root transformation
```{r, message=FALSE}
input_data_na_rm$height_t <- sqrt(input_data_na_rm$height)
input_data_na_rm.scaled$height_t <- standardize(input_data_na_rm$height_t)

hist <- plotNormalHistogram(input_data_na_rm.scaled$height_t, main = "Transformed height")
print(hist)
```




```{r, message=FALSE}
library(MASS)
predictor_old <- input_data_na_rm$sea +1

Box = boxcox(predictor_old ~ 1,   # Transform predictor as a single vector
            lambda = seq(-6,6,0.1))              # Try values -6 to 6 by 0.1
              
Cox = data.frame(Box$x, Box$y)            # Create a data frame with the results

Cox2 = Cox[with(Cox, order(-Cox$Box.y)),] # Order the new data frame by decreasing y

lambda <- Cox2[1, "Box.x"]                 # Extract that lambda
  
print(lambda)

predictor_new <- (predictor_old ^ lambda - 1)/lambda   # Transform the original data

  plotNormalHistogram(predictor_new, main = "sea")
```


# 9. Estimating range of spatial autocorrelation
Spatial autocorrelation in response is estimated based on observations of *Moran's I correlogram* and a *variogram*.

## Moran's I - correlogram
Moran's I results in values from 1 (complete positive spatial autocorrelation) through 0 (no spatial autocorrelation) to -1 (complete negative spatial autocorrelation). 
Correlogram with 100 km range and 10 km bins.

```{r, message=FALSE}
lim.dist <- 5 # range, no.cells (half)
width <- 1 # width, no.cells
M.lags <- moransICorrelogram(input_data_na_rm.scaled$x, input_data_na_rm.scaled$y, 
                             input_data_na_rm.scaled$lc_proportion,lim.dist, width)

```
```{r, echo=FALSE}
ggplot(M.lags,aes(x=x, y=moran)) + geom_line() + xlab("Distance [km]") + ylab("Moran's I") + 
  ggtitle("Correlogram")
```

## Variogram
Variogram with 250 km range and 10 km bins.
```{r, message=FALSE}
library(gstat)
coordinates(input_data_na_rm.scaled)=~x+y

myvariogram <- variogram(lc_proportion~x+y, data=input_data_na_rm.scaled, cutoff=250000, width=10000)  
```

```{r, include=FALSE}
# back to data frame
if (class(input_data_na_rm.scaled) == "SpatialPointsDataFrame") {
  input_data_na_rm.scaled <- as.data.frame(input_data_na_rm.scaled)
}
```

```{r, echo=FALSE}
ggplot(myvariogram,aes(x=dist/1000, y=gamma)) + geom_line() + xlab("Distance [km]") + ylab("Variance") + 
  ggtitle("Variogram")
```


# 10. Explore spatial autocorrelation in residuals from logistic regression
```{r, message=FALSE}
formula.glm <- lc_proportion ~ water_t + curvature + slope_t + bio10 + bio15 + 
                                      bio19 + tpi_10000 + exposure_t + bio02  
fit.glm <- glm(formula = formula.glm, 
               data = input_data_na_rm.scaled,
               family = binomial(link = 'logit')) 
#input_data_na_rm.scaled$glm.res.pearson <- residuals(fit.glm, type="pearson")
#input_data_na_rm.scaled$glm.res.deviance <- residuals(fit.glm, type="deviance")

```

## Correlogram
```{r, message=FALSE}
lim.dist <- 70 # range, no.cells (half)
width <- 10 # width, no.cells
correlog.res.pearson <- moransICorrelogram(input_data_na_rm.scaled$x, input_data_na_rm.scaled$y, 
                                           input_data_na_rm.scaled$glm.res.pearson,lim.dist, width)
correlog.res.deviance <- moransICorrelogram(input_data_na_rm.scaled$x, input_data_na_rm.scaled$y, 
                                           input_data_na_rm.scaled$glm.res.deviance,lim.dist, width)
```

The effective range of spatial autocorrelation of Pearson residuals is approx. 53 km.
The effective range of spatial autocorrelation of deviance residuals is slightly larger (approx. 70 km).
```{r, echo=FALSE}
ggplot(correlog.res.pearson,aes(x=x, y=moran)) + geom_line() + xlab("Distance [km]") + ylab("Moran's I") + 
  ggtitle("Correlogram of Pearson residuals")

ggplot(correlog.res.deviance,aes(x=x, y=moran)) + geom_line() + xlab("Distance [km]") + ylab("Moran's I") + 
  ggtitle("Correlogram of deviance residuals")
```
## Variogram

```{r, message=FALSE}
library(gstat)
coordinates(input_data_na_rm.scaled)=~x+y

variog.res.pearson <- variogram(glm.res.pearson~x+y, data=input_data_na_rm.scaled, cutoff=60000, width=20000)  
#variog.res.deviance <- variogram(glm.res.deviance~x+y, data=input_data_na_rm.scaled, cutoff=100000, width=20000)  
```

```{r, include=FALSE}
# back to data frame
if (class(input_data_na_rm.scaled) == "SpatialPointsDataFrame") {
  input_data_na_rm.scaled <- as.data.frame(input_data_na_rm.scaled)
}
```

```{r, echo=FALSE}
ggplot(variog.res.pearson,aes(x=dist/1000, y=gamma)) + geom_line() + xlab("Distance [km]") + ylab("Variance") + 
  ggtitle("Variogram")
```

## Local spatial autocorrelation
```{r, echo=FALSE}
moran.local.raster <- as.data.frame(moransIlocal(test_data$x, test_data$y, test_data$residuals.pearson.6,10),
                                    xy=TRUE)
ggplot() + geom_tile(data=moran.local.raster[!is.na(moran.local.raster$layer),], aes(x=x, y=y, fill=layer)) + coord_equal() + scale_fill_gradient2(limits=c(-5,5), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0)
```
