---
title: "CAR"
author: "Zofie Cimburova"
date: "6 juli 2017"
output: word_document
---

```{r setup, include=FALSE}
require("knitr")
opts_knit$set(root.dir = "\\\\storage01/zofie.cimburova/My Documents/ecofunc/DATA/sample")
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries
```{r libraries, message=FALSE} 
library(ggplot2)
library(raster)
library(rgdal)  
library(spdep)
library(gstat)
library(rasterVis)
library(CARBayes)
```

# 1. Import observations

Load observation points on the forest line and only keep those observation points situated within a limit vertical distance of 15 m from a forest pixel of local maximum altitude:
```{r import observations} 
forest_line <- read.csv("explanatory_variables.csv")
forest_line <- forest_line[forest_line$lim<=15,]
```

```{r include=FALSE} 
# exclude minimum height
forest_line <- forest_line[forest_line$height > min(forest_line$height),]
# exclude column with ones
forest_line <- subset(forest_line, select = -c(nothing))
# exclude observations too far from others
#sqrt(sum( (A[i,] - B[j,])^2 ))
```

- Split observation points to one dataset for prediction and one dataset for test purposes:
```{r split observations} 
obs <- forest_line[seq(2,nrow(forest_line),2),]
test<- forest_line[seq(1,nrow(forest_line),2),]
```


- Import rasters of explanatory variables.

```{r explanatory, cache = TRUE} 
bio11 <- raster("BIO11_kriged_sample.tif", band=1)
projection(bio11) <- "+proj=utm +zone=33 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"

r.explanatory <- brick(bio11,
                       raster("TPI1010.tif", band=1),
                       raster("lat.tif", band=1), 
                       raster("slope.tif", band=1),
                       raster("srad10.tif", band=1),
                       raster("DEM.tif", band=1),
                       raster("forest_line_10m.tif", band=1))
names(r.explanatory) <- c('BIO11','TPI1010','lat','slope','srad','height','forest_line')
```

# 3. Fit CAR

## check semivariances to find range
```{r ols_semivar, cache = TRUE}
coordinates(obs)<-c('X','Y')

var.mod<-variogram(height~BIO11 + srad + lat + slope + TPI1010,data=obs)
```
```{r ols_variogram, message=FALSE}
plot(var.mod)
```

## 3.1. Calculate 1st order neighbourhood list for CAR model using standard weights.
!!! limit distance d2 needs to be changed so that each observation point has at least one neighbour
```{r car_neigh, cache = TRUE}
neighbours <- dnearneigh(cbind(obs$X, obs$Y), d1 = 0, d2 = 2000)
#neighbours <- knn2nb(knearneigh(cbind(obs$X, obs$Y), k=100))
```

## 3.2. Calculate spatial weights for neighbours lists
```{r car_neigh_weight, cache = TRUE}
listw <- nb2listw(neighbours = neighbours,
                  glist = NULL,
                  style = "W",
                  zero.policy = NULL)
W.mat <- nb2mat(neighbours = neighbours, style = "B")
```

## 3.3. Fit CAR
```{r car_fit, cache = TRUE}
obs <- as.data.frame(obs)
mod.car <- errorsarlm(formula = height~BIO11 + srad + lat + TPI1010, 
                      data = obs,
                      listw = listw)
```

```{r bcar_fit, cache = TRUE}
mod.bcar <- CARBayes::S.CARleroux(formula = height~BIO11 + srad + lat + TPI1010,
                            family='gaussian', 
                            data = obs,
                            W = W.mat, 
                            burnin = 20000, 
                            n.sample = 100000,
                            thin = 10)
```

## 3.4. Predict the whole area
!!! If Error "missing values in newdata", there are missing values in newdata in fact, and need to be excluded (na.action does not help) or replaced or whatever
```{r export, cache = TRUE}
newdata <- as.data.frame(rasterToPoints(r.explanatory))
newdata[is.na(newdata)] <- 0
newdata$predicted_height.car <-predict(mod.car, newdata)


r.height_model.car <- rasterFromXYZ(newdata[,c("x","y","predicted_height.car")])

projection(r.height_model.car) <- "+proj=utm +zone=33"

r.explanatory$forest_line_predict <- r.explanatory$height - r.height_model.car
#writeRaster(r.height_model.car, filename="temp_height_car.tif", format="GTiff", overwrite=TRUE)
```

```{r difference, message=FALSE}

#theme_set(theme_bw())
#gplot(r.forest_line) + geom_tile(aes(fill = value)) +
#          facet_wrap(~ variable) +
#          scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0) +
#          coord_equal()

fl.predict <- levelplot(r.explanatory$forest_line_predict, margin = FALSE, main = "predicted height of forest line", at = unique(c(seq(-150,-10,length=100), seq(-10,0,length=100),seq(0,10,length=100), seq(10,100,length=100))), col.regions = colorRampPalette(c("blue","blue","white","red","red"))(1e3), alpha.regions = 0.35)

fl.obs <- levelplot(r.explanatory$forest_line, margin = FALSE, col.regions = "black" )

print(fl.predict + fl.obs)
```
