---
title: "Logistic regression"
author: "Zofie Cimburova"
date: "28 juni 2017"
output: word_document
---

```{r setup, include=FALSE}
require("knitr")
opts_knit$set(root.dir = "\\\\storage01/zofie.cimburova/My Documents/ecofunc/DATA/sample")
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries
```{r libraries, message=FALSE} 
library(gstat)
library(sp)
library(raster)
library(rgdal)  
library(nnet)
library(spatstat) # for slrm
library(geostatsp) # to convert raster to image
```

# 1. Import observations

- Import rasters of explanatory variables (temperature, TPI, latitude, slope, solar radiation) and response variable (7 - forest, 8 - open land, 9 - forest line):

```{r explanatory} 
bio11 <- raster("BIO11_kriged_sample.tif", band=1)
projection(bio11) <- "+proj=utm +zone=33 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"

r.explanatory <- brick(bio11,
                       raster("TPI1010.tif", band=1),
                       raster("lat.tif", band=1),
                       raster("slope.tif", band=1),
                       raster("srad10.tif", band=1),
                       raster("input_binomial_sample.tif", band=1))
names(r.explanatory) <- c('BIO11','TPI1010','lat','slope','srad10','response')
```


- Generate random observation points (200) across the whole space. Exclude those where some explanatory variable is NaN. Reclass response variable to 0 (open land) and 1 (forest + forest line)
```{r extract}
observations <- sampleRandom(r.explanatory,200, na.rm=TRUE, xy=TRUE)

observations <- as.data.frame(observations)

observations$response[observations$response == 8] <- 0
observations$response[observations$response == 9] <- 0
observations$response[observations$response == 7] <- 1

```


# 2. Fit binomial logistic regression
- Solar radiation excluded due to low significance.
```{r regression}
#mod<- glm(response ~ BIO11 + TPI1010 + lat + slope,family=binomial(link='logit'),data=observations)
#summary(mod)



response.slrm <- as.ppp(observations[,"response"])
bio11.slrm <- as.im(raster("BIO11_kriged_sample.tif", band=1))
tpi1010.slrm <- as.im(raster("TPI1010.tif", band=1))
lat.slrm <- as.im(raster("lat.tif", band=1))
srad.slrm <- as.im(raster("srad10.tif", band=1))

mod.slrm <- slrm(response.slrm ~ bio11.slrm + tpi1010.slrm + lat.slrm + srad.slrm,link = 'logit')

```

## Check spatial autocorrelation in residuals
- apparently clusters of similar values exist
```{r residuals, echo=FALSE}
if (class(observations) != "data.frame") {
  observations <- as.data.frame(observations)
} 

observations$residuals<-data.frame(resids=resid(mod))
```

```{r plot_residuals, echo=FALSE}
coordinates(observations) <- ~x+y
bubble(observations,zcol='resids',main="Residuals in lat-lon space")
```



## Predict the whole area
- Predict probabilities of forest, and then set treshold for forest / open land.
```{r predict}
newdata <- rasterToPoints(r.explanatory)
newdata <- as.data.frame(newdata)

#newdata$predict <- predict(mod,newdata=newdata,type='response')
#newdata$predict10 <- ifelse(newdata$predict > 0.5,1,0)
predict.slrm <- predict(mod.slrm, newdata=list(bio11.slrm, tpi1010.slrm, lat.slrm, srad.slrm), type="probabilities")

#predict.raster <- rasterFromXYZ(newdata)

#misClasificError <- mean(fitted.results != test$Survived)
#print(paste('Accuracy',1-misClasificError))
```
- Plot probabilities and observation points
```{r plot, echo=FALSE}
#coordinates(observations) <- ~x+y

plot(predict.raster$predict)
plot(observations, add=TRUE)
```

- Plot forest / open land and original forest line 
```{r plot10, echo=FALSE}
#coordinates(observations) <- ~x+y

fl <- r.explanatory$response
fl[fl!=9] <- NA

plot(predict.raster$predict10)
plot(fl, col='red', add=TRUE)

```






