---
title: "Bayesian Matérn model fitted with INLA"
output: word_document
---

```{r setup, include=FALSE}
require("knitr")
knitr::opts_chunk$set(echo = TRUE)
source("/home/zofie.cimburova/git/oekokart/ecofunc_helpful_functions.R")
opts_knit$set(root.dir = "/home/zofie.cimburova/ECOFUNC/DATA/OBSERVATIONS")
```

```{r, include=FALSE}
library(ggplot2)
```

Predict probability of success in a cell from environmental variables.

### Problem I encountered:
Whilst running the model with fixed effects only always gives indentical results, including random effects causes the results to differ. Globally (expressed as mean absolute difference), the differences are not very big (at third decimal place), but locally may be much larger, causing a pixel to be classified as forest in one run and open land in second run.


Subset of data to test on:
```{r, include=FALSE}
# subset for testing
#test_inp_data <- input_data_na_rm.scaled[input_data_na_rm.scaled$x < 400000 & # x (-75500, 1156500)
#                                     input_data_na_rm.scaled$x >=  0 & 
#                                     input_data_na_rm.scaled$y <  7100000 & # y (6451500, 7939500)
#                                     input_data_na_rm.scaled$y >= 6700000,]
test_inp_data <- input_data_na_rm.scaled

# number of rows and columns
#nrows=ncols=400
#nrows=1808
#ncols=1414
nrows=nrows.test=497
ncols=ncols.test=412

#test_inp_data$row <- test_inp_data$row - min(test_inp_data$row) + 1
#test_inp_data$col <- test_inp_data$col - min(test_inp_data$col) + 1
```

# 0. Divide data into one dataset to fit and another dataset to validate
Regular grid.
Every 3rd pixel.
```{r, message=FALSE}
# only even rows and even columns for fitting
#test_data <- test_inp_data[(test_inp_data$row%%3 == 0 & test_inp_data$col%%3 == 0),]
test_data <- test_inp_data

# rest for validation
#validate_data <- test_inp_data[(test_inp_data$row%%3 != 0 | test_inp_data$col%%3 != 0),]

# change row and column number in test data (squeeze)
#test_data$row <- test_data$row/3
#test_data$col <- test_data$col/3

# assign INLA ID
#nrows.test = floor(nrows/3)
#ncols.test = floor(ncols/3)
#test_data$ID <- inla.lattice2node(test_data$row, test_data$col, nrows.test, ncols.test)

# order data by ID
test_data <- test_data[order(test_data$ID),]

```
```{r, echo=FALSE}
ggplot() + geom_tile(data=test_inp_data, aes(x=x, y=y, fill=lc)) + coord_equal() + ggtitle("Full data")
#ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=lc)) + coord_equal() + ggtitle("Data for fitting")
#ggplot() + geom_tile(data=validate_data, aes(x=x, y=y, fill=lc)) + coord_equal() + ggtitle("Validation data")
```

# 1. Shapes of Matern covariance function
Effective range of spatial autocorrelation in proportion of forest / open land: 125 km
Effective range of spatial autocorrelation in pearson residuals of GLM: 53 km
Effective range of spatial autocorrelation in deviance residuals of GLM: 70 km

```{r, message=FALSE}
range=53/3
nu=1
```


# 2. Fit a model with random and fixed effects

```{r, message=FALSE}
require(INLA)
models_complete <- list()
#models_400 <- list()

formula1 <- lc ~ water + curvature + slope + bio10 + bio15 + bio19 + tpi_10000 + exposure + bio02 + height + 
                            f(ID, model='matern2d',
                              nrow=nrows.test, ncol=nrows.test, nu=nu,
                              hyper=list(range=list(initial=log(range), fixed=TRUE),
                                         prec=list(initial=log(precision), fixed=TRUE))) 
formula2 <- lc ~ water_t + slope_t + bio10 + bio15 + bio19 + tpi_10000 + exposure_t + bio02 + height_t + 
                            f(ID, model='matern2d',
                              nrow=nrows.test, ncol=nrows.test, nu=nu,
                              hyper=list(range=list(initial=log(range), fixed=TRUE),
                                         prec=list(initial=log(precision), fixed=TRUE))) 
formula3 <- lc ~ water_t + slope_t + bio10 + bio15 + bio19 + tpi_10000 + exposure_t + bio02 + height_t + solrad_jan +
                            f(ID, model='matern2d',
                              nrow=nrows.test, ncol=nrows.test, nu=nu,
                              hyper=list(range=list(initial=log(range), fixed=TRUE),
                                         prec=list(initial=log(precision), fixed=TRUE))) 

formulas = c(formula1)

i = 1
#for (precision in c(0.01,0.1,1,10,100,1000,10000,100000)) {
for (formula in formulas) {
  precision = 10000
  
  
  matern <- inla(formula=formula,
                  data=test_data,
                  family='binomial',
                  Ntrials=1,
                  verbose=F,
                  control.inla=list(tolerance=1e-6), # control.compute=list(dic=TRUE, waic=TRUE), 
                  control.fixed=list(prec.intercept=0.001), # can be omitted 
                  control.predictor=list(link=1),
                  num.threads = 11)
  
  # save model  
  models_complete[[i]] <- matern
  i = i+1
}
```

Identical model without random effect for comparison:
```{r, message=FALSE}
formula.matern.9 <- nsuccess ~ water_t + curvature + slope_t + bio10 + bio15 + bio19 + tpi_10000 + exposure_t + bio02 + height_t
                            
matern <- inla(formula=formula.matern.9,
                  data=test_data,
                  family='binomial',
                  Ntrials=ntrials,
                  verbose=F,
                  control.inla=list(tolerance=1e-6),
                  control.compute=list(dic=TRUE, waic=TRUE),
                  control.fixed=list(prec.intercept=0.001), # can be omitted
                  control.predictor=list(link=1),
                  num.threads = 12) 

# save model  
models_complete[[i]] <- matern
i = i+1
```

# 3. Summarize results:
```{r, include=FALSE}
# fitted values (DO NOT SHUFFLE INPUT DATA FRAME!!!)
test_data <- append.fitted(test_data, models_complete[[1]]$summary.fitted.values, "fitted.values.1")
#test_data <- append.fitted(test_data, models_400[[2]]$summary.fitted.values, "fitted.values.2")
#test_data <- append.fitted(test_data, models_400[[3]]$summary.fitted.values, "fitted.values.3")
#test_data <- append.fitted(test_data, models_complete[[4]]$summary.fitted.values, "fitted.values.4")
#test_data <- append.fitted(test_data, models_complete[[5]]$summary.fitted.values, "fitted.values.5")
#test_data <- append.fitted(test_data, models_complete[[6]]$summary.fitted.values, "fitted.values.6")
#test_data <- append.fitted(test_data, models_complete[[7]]$summary.fitted.values, "fitted.values.7")
#test_data <- append.fitted(test_data, models_complete[[8]]$summary.fitted.values, "fitted.values.8")
#test_data <- append.fitted(test_data, models_complete[[9]]$summary.fitted.values, "fitted.values.9")

# absolute residuals (p)
test_data$residuals.abs.1 <- test_data$lc - test_data$fitted.values.1
#test_data$residuals.abs.2 <- test_data$lc_proportion - test_data$fitted.values.2
#test_data$residuals.abs.3 <- test_data$lc_proportion - test_data$fitted.values.3
#test_data$residuals.abs.4 <- test_data$lc_proportion - test_data$fitted.values.4
#test_data$residuals.abs.5 <- test_data$lc_proportion - test_data$fitted.values.5
#test_data$residuals.abs.6 <- test_data$lc_proportion - test_data$fitted.values.6
#test_data$residuals.abs.7 <- test_data$lc_proportion - test_data$fitted.values.7
#test_data$residuals.abs.8 <- test_data$lc_proportion - test_data$fitted.values.8
#test_data$residuals.abs.9 <- test_data$lc_proportion - test_data$fitted.values.9

# pearson residuals (logit(p))
#test_data$residuals.pearson.1 <- residuals.pearson(test_data$lc,test_data$fitted.values.1)
#test_data$residuals.pearson.2 <- residuals.pearson(test_data$lc_proportion,test_data$fitted.values.2)
#test_data$residuals.pearson.3 <- residuals.pearson(test_data$lc_proportion,test_data$fitted.values.3)
#test_data$residuals.pearson.4 <- residuals.pearson(test_data$lc_proportion,test_data$fitted.values.4)
#test_data$residuals.pearson.5 <- residuals.pearson(test_data$lc_proportion,test_data$fitted.values.5)
#test_data$residuals.pearson.6 <- residuals.pearson(test_data$lc_proportion,test_data$fitted.values.6)
#test_data$residuals.pearson.7 <- residuals.pearson(test_data$lc_proportion,test_data$fitted.values.7)
#test_data$residuals.pearson.8 <- residuals.pearson(test_data$lc_proportion,test_data$fitted.values.8)
#test_data$residuals.pearson.9 <- residuals.pearson(test_data$lc_proportion,test_data$fitted.values.9)

# random effects (p)
test_data <- append.random(test_data, models_complete[[1]]$summary.random$ID, "random.effect.1")
#test_data <- append.random(test_data, models_400[[2]]$summary.random$ID, "random.effect.2")
#test_data <- append.random(test_data, models_400[[3]]$summary.random$ID, "random.effect.3")
#test_data <- append.random(test_data, models_complete[[4]]$summary.random$ID, "random.effect.4")
#test_data <- append.random(test_data, models_complete[[5]]$summary.random$ID, "random.effect.5")
#test_data <- append.random(test_data, models_complete[[6]]$summary.random$ID, "random.effect.6")
#test_data <- append.random(test_data, models_complete[[7]]$summary.random$ID, "random.effect.7")
#test_data <- append.random(test_data, models_complete[[8]]$summary.random$ID, "random.effect.8")

# fixed effects (p)
observed.values <- test_data[,c("water","curvature","slope","bio10",
                                "bio15","bio19","tpi_10000","exposure","bio02","height")]
test_data$fixed.effect.1 <- fixed.effects(models_complete[[1]]$summary.fixed$mean, observed.values)
#observed.values <- test_data[,c("water_t","slope_t","bio10",
#                                "bio15","bio19","tpi_10000","exposure_t","bio02","height_t")]
#test_data$fixed.effect.2 <- fixed.effects(models_400[[2]]$summary.fixed$mean, observed.values)
#observed.values <- test_data[,c("water_t","slope_t","bio10",
#                                "bio15","bio19","tpi_10000","exposure_t","bio02","height_t","solrad_jan")]
#test_data$fixed.effect.3 <- fixed.effects(models_400[[3]]$summary.fixed$mean, observed.values)
#test_data$fixed.effect.4 <- fixed.effects(models_complete[[4]]$summary.fixed$mean, observed.values)
#test_data$fixed.effect.5 <- fixed.effects(models_complete[[5]]$summary.fixed$mean, observed.values)
#test_data$fixed.effect.6 <- fixed.effects(models_complete[[6]]$summary.fixed$mean, observed.values)
#test_data$fixed.effect.7 <- fixed.effects(models_complete[[7]]$summary.fixed$mean, observed.values)
#test_data$fixed.effect.8 <- fixed.effects(models_complete[[8]]$summary.fixed$mean, observed.values)
#test_data$fixed.effect.9 <- fixed.effects(models_complete[[9]]$summary.fixed$mean, observed.values)

# absolute residuals of fixed effect (p)
test_data$residuals.fixed.1 <- test_data$lc_proportion - test_data$fixed.effect.1
#test_data$residuals.fixed.2 <- test_data$lc_proportion - test_data$fixed.effect.2
#test_data$residuals.fixed.3 <- test_data$lc_proportion - test_data$fixed.effect.3
#test_data$residuals.fixed.4 <- test_data$lc_proportion - test_data$fixed.effect.4
#test_data$residuals.fixed.5 <- test_data$lc_proportion - test_data$fixed.effect.5
#test_data$residuals.fixed.6 <- test_data$lc_proportion - test_data$fixed.effect.6
#test_data$residuals.fixed.7 <- test_data$lc_proportion - test_data$fixed.effect.7
#test_data$residuals.fixed.8 <- test_data$lc_proportion - test_data$fixed.effect.8
#test_data$residuals.fixed.9 <- test_data$lc_proportion - test_data$fixed.effect.9
```

## Conclusion 1: Influence of precision on coefficients of fixed effects
```{r, include=FALSE}
library("reshape2")
coefficients.summary <- as.data.frame(rbind(models_complete[[1]]$summary.fixed$mean,
                                            models_complete[[2]]$summary.fixed$mean,
                                            models_complete[[3]]$summary.fixed$mean,
                                            models_complete[[4]]$summary.fixed$mean,
                                            models_complete[[5]]$summary.fixed$mean,
                                            models_complete[[6]]$summary.fixed$mean,
                                            models_complete[[7]]$summary.fixed$mean,
                                            models_complete[[8]]$summary.fixed$mean))

coefficients.summary.glm <- as.data.frame(rbind(models_complete[[9]]$summary.fixed$mean))

colnames(coefficients.summary) <- c("Intercept","water_t","curvature","slope_t","bio10",
                                "bio15","bio19","tpi_10000","exposure_t","bio02","height_t")
colnames(coefficients.summary.glm) <- c("Intercept","water_t","curvature","slope_t","bio10",
                                "bio15","bio19","tpi_10000","exposure_t","bio02","height_t")

coefficients.summary$precision <- c(0.01,0.1,1,10,100,1000,10000,100000)
coefficients.summary.glm$precision <- c(110000)

coefficients.summary.long <- melt(coefficients.summary,id="precision")
coefficients.summary.glm.long <- melt(coefficients.summary.glm,id="precision")
```

```{r, echo=FALSE}
ggplot() + 
  geom_line(data=coefficients.summary.long, aes(x=precision, y=value, colour=variable)) +
  geom_point(data=coefficients.summary.glm.long, aes(x=precision, y=value, colour=variable)) + 
  scale_color_manual(values=c("gray36", "dodgerblue4","gold", "darkcyan","firebrick2", 
                              "darkorange2","forestgreen", "darkolivegreen3","hotpink", "tomato2","red"))
```

## Conclusion 2: Influence of precision on fixed effects, random effects and residuals
```{r, echo=FALSE, fig.height = 10, fig.width = 10}
style = theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank())

ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=lc_proportion)) + coord_equal() + ggtitle("original data") +  scale_fill_gradient(limits=c(0,1), low = "bisque3", high = "darkgreen")

# fitted values
a1 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=fitted.values.1)) + coord_equal() + ggtitle("fitted values 1") +  scale_fill_gradient(limits=c(0,1), low = "bisque3", high = "darkgreen") + style
a2 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=fitted.values.2)) + coord_equal() + ggtitle("fitted values 2") +  scale_fill_gradient(limits=c(0,1), low = "bisque3", high = "darkgreen") + style
a3 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=fitted.values.3)) + coord_equal() + ggtitle("fitted values 3") +  scale_fill_gradient(limits=c(0,1), low = "bisque3", high = "darkgreen") + style
#a4 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=fitted.values.4)) + coord_equal() + ggtitle("fitted values 4") +  scale_fill_gradient(limits=c(0,1), low = "bisque3", high = "darkgreen") + style
#a5 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=fitted.values.5)) + coord_equal() + ggtitle("fitted values 5") +  scale_fill_gradient(limits=c(0,1), low = "bisque3", high = "darkgreen") + style
#a6 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=fitted.values.6)) + coord_equal() + ggtitle("fitted values 6") +  scale_fill_gradient(limits=c(0,1), low = "bisque3", high = "darkgreen") + style
#a7 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=fitted.values.7)) + coord_equal() + ggtitle("fitted values 7") +  scale_fill_gradient(limits=c(0,1), low = "bisque3", high = "darkgreen") + style
#a8 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=fitted.values.8)) + coord_equal() + ggtitle("fitted values 8") +  scale_fill_gradient(limits=c(0,1), low = "bisque3", high = "darkgreen") + style
#a9 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=fitted.values.9)) + coord_equal() + ggtitle("fitted values 9") +  scale_fill_gradient(limits=c(0,1), low = "bisque3", high = "darkgreen") + style

multiplot(a1,a2,a3,cols=2)#,a4,a5,a6,a7,a8,a9,cols=2)

# fixed effect
b1 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=fixed.effect.1)) + coord_equal() + ggtitle("fixed effect 1") +  scale_fill_gradient(limits=c(0,1), low = "bisque3", high = "darkgreen") + style
b2 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=fixed.effect.2)) + coord_equal() + ggtitle("fixed effect 2") +  scale_fill_gradient(limits=c(0,1), low = "bisque3", high = "darkgreen") + style
b3 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=fixed.effect.3)) + coord_equal() + ggtitle("fixed effect 3") +  scale_fill_gradient(limits=c(0,1), low = "bisque3", high = "darkgreen") + style
#b4 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=logit(fixed.effect.4))) + coord_equal() + ggtitle("fixed effect 4") +  scale_fill_gradient2(limits=c(-10,10), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
#b5 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=logit(fixed.effect.5))) + coord_equal() + ggtitle("fixed effect 5") +  scale_fill_gradient2(limits=c(-10,10), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
#b6 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=logit(fixed.effect.6))) + coord_equal() + ggtitle("fixed effect 6") +  scale_fill_gradient2(limits=c(-10,10), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
#b7 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=logit(fixed.effect.7))) + coord_equal() + ggtitle("fixed effect 7") +  scale_fill_gradient2(limits=c(-10,10), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
#b8 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=logit(fixed.effect.8))) + coord_equal() + ggtitle("fixed effect 8") +  scale_fill_gradient2(limits=c(-10,10), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
#b9 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=logit(fixed.effect.9))) + coord_equal() + ggtitle("fixed effect 9") +  scale_fill_gradient2(limits=c(-10,10), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style

multiplot(b1,b2,b3,cols=2)#,b4,b5,b6,b7,b8,b9,cols=2)

# residuals
c1 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=residuals.abs.1)) + coord_equal() + ggtitle("absolute residuals 1") + scale_fill_gradient2(limits=c(-1,1), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
c2 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=residuals.abs.2)) + coord_equal() + ggtitle("absolute residuals 2") + scale_fill_gradient2(limits=c(-1,1), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
c3 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=residuals.abs.3)) + coord_equal() + ggtitle("absolute residuals 3") + scale_fill_gradient2(limits=c(-1,1), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
#c4 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=residuals.abs.4)) + coord_equal() + ggtitle("absolute residuals 4") + scale_fill_gradient2(limits=c(-1,1), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
#c5 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=residuals.abs.5)) + coord_equal() + ggtitle("absolute residuals 5") + scale_fill_gradient2(limits=c(-1,1), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
#c6 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=residuals.abs.6)) + coord_equal() + ggtitle("absolute residuals 6") + scale_fill_gradient2(limits=c(-1,1), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
#c7 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=residuals.abs.7)) + coord_equal() + ggtitle("absolute residuals 7") + scale_fill_gradient2(limits=c(-1,1), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
#c8 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=residuals.abs.8)) + coord_equal() + ggtitle("absolute residuals 8") + scale_fill_gradient2(limits=c(-1,1), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
#c9 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=residuals.abs.9)) + coord_equal() + ggtitle("absolute residuals 9") + scale_fill_gradient2(limits=c(-1,1), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style

multiplot(c1,c2,c3,cols=2)#,c4,c5,c6,c7,c8,c9,cols=2)

# random effect
d1 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=logit(random.effect.1))) + coord_equal() + ggtitle("random effect 1") +  scale_fill_gradient2(limits=c(-0.04,0.04), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
d2 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=logit(random.effect.2))) + coord_equal() + ggtitle("random effect 2") +  scale_fill_gradient2(limits=c(-10,10), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
d3 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=logit(random.effect.3))) + coord_equal() + ggtitle("random effect 3") +  scale_fill_gradient2(limits=c(-10,10), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
#d4 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=logit(random.effect.4))) + coord_equal() + ggtitle("random effect 4") +  scale_fill_gradient2(limits=c(-10,10), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
#d5 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=logit(random.effect.5))) + coord_equal() + ggtitle("random effect 5") +  scale_fill_gradient2(limits=c(-10,10), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
#d6 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=logit(random.effect.6))) + coord_equal() + ggtitle("random effect 6") +  scale_fill_gradient2(limits=c(-10,10), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
#d7 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=logit(random.effect.7))) + coord_equal() + ggtitle("random effect 7") +  scale_fill_gradient2(limits=c(-10,10), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
#d8 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=logit(random.effect.8))) + coord_equal() + ggtitle("random effect 8") +  scale_fill_gradient2(limits=c(-10,10), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style

multiplot(d1,d2,d3,cols=2)#,d4,d5,d6,d7,d8,cols=2)

# absolute residuals of fixed effect
e1 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=residuals.fixed.1)) + coord_equal() + ggtitle("absolute residuals of fixed effect 1") +  scale_fill_gradient2(limits=c(-1,1), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
e2 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=residuals.fixed.2)) + coord_equal() + ggtitle("absolute residuals of fixed effect 2") +  scale_fill_gradient2(limits=c(-1,1), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
e3 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=residuals.fixed.3)) + coord_equal() + ggtitle("absolute residuals of fixed effect 3") +  scale_fill_gradient2(limits=c(-1,1), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
#e4 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=residuals.fixed.4)) + coord_equal() + ggtitle("absolute residuals of fixed effect 4") +  scale_fill_gradient2(limits=c(-1,1), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
#e5 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=residuals.fixed.5)) + coord_equal() + ggtitle("absolute residuals of fixed effect 5") +  scale_fill_gradient2(limits=c(-1,1), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
#e6 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=residuals.fixed.6)) + coord_equal() + ggtitle("absolute residuals of fixed effect 6") +  scale_fill_gradient2(limits=c(-1,1), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
#e7 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=residuals.fixed.7)) + coord_equal() + ggtitle("absolute residuals of fixed effect 7") +  scale_fill_gradient2(limits=c(-1,1), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
#e8 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=residuals.fixed.8)) + coord_equal() + ggtitle("absolute residuals of fixed effect 8") +  scale_fill_gradient2(limits=c(-1,1), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style
#e9 <- ggplot() + geom_tile(data=test_data, aes(x=x, y=y, fill=residuals.fixed.9)) + coord_equal() + ggtitle("absolute residuals of fixed effect 9") +  scale_fill_gradient2(limits=c(-1,1), low = "dodgerblue2", mid = "cornsilk1", high = "red2", midpoint = 0) + style

multiplot(e1,e2,e3,cols=2)#,e4,e5,e6,e7,e8,e9,cols=2)
```


## Conclusion 3: Influence of precision on spatial autocorrelation in Pearson residuals
```{r, message=FALSE}
  # choose random samples
  sample.moran.I <- test_data[sample(nrow(test_data), 10000),c("x","y","residuals.pearson.1","residuals.pearson.2",
                                                              "residuals.pearson.3","residuals.pearson.4",
                                                              "residuals.pearson.5","residuals.pearson.6") ]

  print(morans.I.ape(sample.moran.I,1000,"residuals.pearson.1"))
  print(morans.I.ape(sample.moran.I,1000,"residuals.pearson.2"))
  print(morans.I.ape(sample.moran.I,1000,"residuals.pearson.3"))
  print(morans.I.ape(sample.moran.I,1000,"residuals.pearson.4"))
  print(morans.I.ape(sample.moran.I,1000,"residuals.pearson.5"))
  print(morans.I.ape(sample.moran.I,1000,"residuals.pearson.6"))
```


```{r, echo=FALSE}
sa.summary <- data.frame(precision <- c(10,100,1000,10000,100000),
                         sa50 <- c(0.00618154552186201,0.013143425826186,0.0255525641332435,0.0339726111956901,0.0438487773441474),
                         sa10 <- c(0.00623831723683146,0.00726044290065922,0.0121028693862007,0.0394502005779933,0.0896333635937407))
ggplot() + geom_line(data=sa.summary, aes(x=precision, y=sa50, color='blue')) +
  geom_line(data=sa.summary, aes(x=precision, y=sa10, color='red')) + 
  geom_point(aes(x=110000,y=0.20309925773823,color='blue')) + 
  geom_point(aes(x=110000,y=0.44197766618071,color='red'))
```

## 5. Influence of precision and range on various parametres
```{r, include=FALSE}
criteria.precision <- data.frame(precision = seq(0.05,20,0.05),
                                 DIC = DIC,
                                 WAIC = WAIC,
                                 pD = pD,
                                 N_pD = N_pD)
criteria.range <- data.frame(range = seq(1,200,1),
                             DIC = DIC,
                             WAIC = WAIC,
                             pD = pD,
                             N_pD = N_pD)


ggplot(criteria.precision) + geom_line(aes(x=precision, y=DIC)) + xlab("Precision") + ylab("DIC") 
ggplot(criteria.precision) + geom_line(aes(x=precision, y=WAIC)) + xlab("Precision") + ylab("WAIC") 
ggplot(criteria.precision) + geom_line(aes(x=precision, y=pD)) + xlab("Precision") + ylab("Number of estimated parametres") 
ggplot(criteria.precision) + geom_line(aes(x=precision, y=N_pD)) + xlab("Precision") + ylab("Number of effective replicates") 

ggplot(criteria.range) + geom_line(aes(x=range, y=DIC)) + xlab("Range") + ylab("DIC") 
ggplot(criteria.range) + geom_line(aes(x=range, y=WAIC)) + xlab("Range") + ylab("WAIC") 
ggplot(criteria.range) + geom_line(aes(x=range, y=pD)) + xlab("Range") + ylab("Number of estimated parametres") 
ggplot(criteria.range) + geom_line(aes(x=range, y=N_pD)) + xlab("Range") + ylab("Number of effective replicates") 

```


## 4. Export fixed effects
```{r, include=FALSE}
output_data <- cbind(rownames(models_complete[[1]]$summary.fixed), 
                     models_complete[[1]]$summary.fixed$mean,
                     c(0,colMeans(input_data_na_rm[,c("water_t","curvature","slope_t",
                                                  "bio10","bio15","bio19","tpi_10000",
                                                  "exposure_t","bio02","height_t")])),
                     c(0,apply(input_data_na_rm[,c("water_t","curvature","slope_t",
                                               "bio10","bio15","bio19","tpi_10000",
                                               "exposure_t","bio02","height_t")], 2, sd)))

write.table(output_data,
            file="/home/zofie.cimburova/ECOFUNC/DATA/OBSERVATIONS/fixed_effects_coefficients.csv",sep=',',quote=FALSE)
```

## 5. Export random effects
```{r, include=FALSE}
raster <- rasterFromXYZ(cbind(test_data$x,test_data$y,logit(test_data$random.effect.1)))
crs(raster) <- "+proj=utm +zone=33 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
writeRaster(raster, "/home/zofie.cimburova/ECOFUNC/DATA/OBSERVATIONS/random_effect", format = "GTiff", overwrite=TRUE)
```






